Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-03 PandaEvolution.tex

Currently, very few HEP applications could benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks of ATLAS workflows can
be offloaded from the Grid to Titan's large amount of cores. Further, when HEP
tasks can be partitioned to independent jobs, Titan worker nodes can be used to
execute up to 16 concurrent payloads, one for each available core. Given these
constraints and challenges, the type of task most suitable for execution at the
moment on Titan is Monte Carlo detector simulation. This type of task is mostly
computational-intensive, requiring less than 2GB of RAM at runtime and with
small input data requirements. Detector simulation tasks in ATLAS are performed
via AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 simulation toolkit~\cite{agostinelli2003geant4}. These tasks account for
$\approx$ 60\% of all the jobs on WLCG, making them a primary candidate for
offloading.

>>>>

Currently, very few HEP applications can benefit from Titan's GPUs but some
computationally-intensive and non memory-intensive tasks of ATLAS workflows can
be offloaded from the Grid to Titan's large amount of cores. Further, when HEP
tasks can be partitioned to independent jobs, Titan worker nodes can be used to
execute up to 16 concurrent payloads, one for each available core. Given these
constraints and challenges, the type of task most suitable for execution at the
moment on Titan is Monte Carlo detector simulation. This type of task is mostly
computational-intensive, requiring less than 2GB of RAM at runtime and with
small input data requirements. Detector simulation tasks in ATLAS are performed
via AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 simulation toolkit~\cite{agostinelli2003geant4}. These tasks account for
$\approx$ 60\% of all the jobs on WLCG, making them a primary candidate for
offloading.

### 2017-04-03 nge.tex

\jhanote{not sure MPI script is optimal description or ambiguous.}\mtnote{Is
what we used in the previous section(s) to designate what PanDA Broker submits
to Titan's Torque batch system. Job is already used for what is pulled from
PanDA server.} Multiple generations would also relax two assumptions of the
current execution model: knowing the number of simulations before submitting the
MPI script, and having a fixed number of events per simulation (100 at the
moment). Pilots would enable the scheduling of simulations independently from
whether they were available at the moment of submitting the pilot. Further,
simulations with a varying number of events could be scheduled on a pilot,
depending on the amount of remaining walltime and the distribution of execution
time per event, as shown in \S\ref{ssec:panda_titan},
Figure~\ref{fig:100event-distrib}. These capabilities would increase the
efficiency of the PanDA Broker when there is a large difference between the
number of cores and walltime.

@@

NGE exposes an API for the application  \jhanote{workload management?} layer to
describe workloads (Fig.~\ref{fig:arch-overview}, green squares) and pilots
(Fig.~\ref{fig:arch-overview}, red circles), and to instantiate a PilotManager
and a UnitManager.   The PilotManager submits pilots to Titan's Torque
\jhanote{for consistency should we say Moab?}\mtnote{Moab = Scheduler; Torque =
Batch system.} batch system via SAGA API (Fig.~\ref{fig:arch-overview}, dash
arrow), as done by PanDA Broker to submit MPI scripts. Once scheduled, the
pilot's Agent is bootstrapped on Titan's MOM node and \jhanote{the Agent's
Executors on the} the Agent's Executors on the worker nodes, and the
UnitManager schedules units to the Agent's Scheduler (Fig.~\ref{fig:arch-
overview}, solid arrow). The Agent's Scheduler schedules the units on one or
more Agent's Executor for execution. The Agent's executors can manage one or
more worker nodes, depending on performance evaluations. The UnitManager and the
Agent communicate via a database that is instantiated on one of Titan's DTN so
as to be reachable by both modules.

>>>>

\jhanote{not sure MPI script is optimal description or ambiguous.}\mtnote{Is
what we used in the previous section(s) to designate what PanDA Broker submits
to Titan's PBS batch system. Job is already used for what is pulled from
PanDA server.} Multiple generations would also relax two assumptions of the
current execution model: knowing the number of simulations before submitting the
MPI script, and having a fixed number of events per simulation (100 at the
moment). Pilots would enable the scheduling of simulations independently from
whether they were available at the moment of submitting the pilot. Further,
simulations with a varying number of events could be scheduled on a pilot,
depending on the amount of remaining walltime and the distribution of execution
time per event, as shown in \S\ref{ssec:panda_titan},
Figure~\ref{fig:100event-distrib}. These capabilities would increase the
efficiency of the PanDA Broker when there is a large difference between the
number of cores and walltime.

@@

NGE exposes an API for the application  \jhanote{workload management?} layer to
describe workloads (Fig.~\ref{fig:arch-overview}, green squares) and pilots
(Fig.~\ref{fig:arch-overview}, red circles), and to instantiate a PilotManager
and a UnitManager.   The PilotManager submits pilots to Titan's PBS \jhanote{for
consistency should we say Moab?}\mtnote{Moab = Scheduler; Torque = Batch system.
I used PBS Batch System because it makes easier to speak about both the system
and the script submitted to it.} batch system via SAGA API
(Fig.~\ref{fig:arch-overview}, dash arrow), as done by PanDA Broker to submit
MPI scripts. Once scheduled, the pilot's Agent is bootstrapped on Titan's MOM
node and \jhanote{the Agent's Executors on the} the Agent's Executors on the
worker nodes, and the UnitManager schedules units to the Agent's Scheduler
(Fig.~\ref{fig:arch- overview}, solid arrow). The Agent's Scheduler schedules
the units on one or more Agent's Executor for execution. The Agent's executors
can manage one or more worker nodes, depending on performance evaluations. The
UnitManager and the Agent communicate via a database that is instantiated on one
of Titan's DTN so as to be reachable by both modules.
